Forecasting Gamma-ray Bursts using Gravitational Waves

### 2018-02-23 Paper/Manuscript.tex
First commit

\documentclass[11pt,a4paper]{article}

\usepackage{multicol}
\usepackage{wrapfig}
\usepackage{graphicx}
\usepackage{color}
\usepackage{amsfonts,amssymb,amsmath}
     % use Times fonts if available on your TeX system
\usepackage{epsfig}
%\usepackage{empheq}
\usepackage{subfigure}
\usepackage{hyperref}
\usepackage{rotating}
\usepackage[small,compact]{titlesec}
\usepackage{fancyhdr}
\usepackage{times}
\usepackage{multirow}
\usepackage{sidecap}
\usepackage{eurosym}
\usepackage{setspace}
\usepackage{caption}
\usepackage[usestackEOL]{stackengine} % package used to write several lines in one box
%\usepackage{subcaption}

\topmargin-14mm
\textheight250mm
\oddsidemargin1mm
\textwidth160mm

\newcommand{\la}{\langle}
\newcommand{\ra}{\rangle}
\renewcommand{\c}{\cos\theta}
\newcommand{\s}{\sin}
\renewcommand{\t}{\theta}
\renewcommand{\b}{\bar}
\newcommand{\f}{\frac}
\newcommand{\df}{\dfrac}
\newcommand{\pd}{\partial}
\newcommand{\bt}{\beta}
\newcommand{\D}{\Delta}
\newcommand{\Si}{\Sigma}
\newcommand{\p}{\rho}
\renewcommand{\L}{\mathcal{L}}
\newcommand{\ph}{\phi}
\newcommand{\g}{\gamma}
\newcommand{\nn}{\nonumber}
\newcommand{\al}{\alpha}
\newcommand{\La}{\Lambda}
\newcommand{\el}{\ell}
\newcommand{\h}{\hat}
\newcommand{\mrm}{\mathrm}
\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\ba}{\begin{align}}
\newcommand{\ea}{\end{align}}
\newcommand{\bi}{\begin{itemize}}
\newcommand{\ei}{\end{itemize}}
\newcommand{\bef}{\begin{frame}}
\newcommand{\ef}{\end{frame}}
\newcommand{\bs}{\begin{small}}
\newcommand{\es}{\end{small}}
\newcommand{\ord}{\mathcal{O}}
\newcommand{\Lim}[1]{\raisebox{0.5ex}{\scalebox{0.8}{$\displaystyle \lim_{#1}\;$}}}
\newcommand\T{\rule{0pt}{2.6ex}}       % Top strut for vertical spacing in tables
\newcommand\B{\rule[-1.2ex]{0pt}{0pt}} % Bottom strut

\begin{document}
{\bf Title: Forecasting Gamma-ray Bursts using Gravitational Waves}
\section{Abstract}
We explore the intriguing possibility of employing future gravitational-wave interferometers to detect the inspiral of compact binaries early enough to alert the electromagnetic observatories so that a gamma-ray burst (GRB) can be observed in its entirety from its very start.
We quantify the ability to predict a GRB by computing the time a canonical binary neutron star (BNS) system takes to inspiral from a certain threshold frequency to its final merger. We define this threshold frequency to be the gravitational-wave frequency at which the RMS-averaged strain due to the BNS system accumulates a signal-to-noise (SNR) ratio of 15. For our computation of advance warning times, we consider BNS systems at luminosity distances of (i) $D\lesssim200\,$Mpc in a three-interferometer network with Advance LIGO's design sensitivity and (ii) $D \lesssim 1000\,$Mpc in a similar network, but with the Einstein Telescope's B configuration
sensitivity. We find in the case of that Advanced LIGO we may get a few minutes of warning time if we get lucky whereas the Einstein Telescope will provide us
with warning times as long as a few hours.

We also consider binary black hole neutron star systems with the hopes of observing a tidal disruption event in the future. However, our findings indicate that
witnessing such an event will be very unlikely even after the Einstein Telescope becomes operational.

@@

@@

Let us now turn our attention to the GWs generated by the inspiral.
The GWs themselves actually are tensor distortions $h_{\mu\nu}$ of the flat Minkowski
spacetime propagating at the speed of light.
In more technical terms, they satisfy $\Box h_{\mu\nu} = 0$, i.e., the flat spacetime sourceless wave equation
meaning that we are considering the solutions far from the GW source, in the so-called wave zone defined by the condition $ c/f \ll D$.
In a suitable gauge, such as the commonly used TT (transverse-traceless) gauge, it can be shown that
there exist only two physical degrees of freedom which manifest themselves
as two independent polarization amplitudes $h_+$ and $h_\times$. %when interacting with masses away from the source.
For a circular binary at a distance $D$ these states read
%
\begin{align}
 h_+(t) &= \f{4}{D}\left(\f{G M_c}{c^2}\right)^{5/3}\, \left(\f{\pi f(t)}{c}\right)^{2/3}\, \left(\f{1+\cos^2\iota}{2}\right)\, \cos[\Phi_\text{N}(t)],\label{eq:hplus_TD}\\
 h_\times(t) & = \f{4}{D}\left(\f{G M_c}{c^2}\right)^{5/3}\, \left(\f{\pi f(t)}{c}\right)^{2/3}\,\cos\iota \sin[\Phi_\text{N}(t)]\label{eq:hcross_TD},
\end{align}
%
%
where $\iota=\cos^{-1}(\mathbf{\hat{n}}\cdot{\mathbf{\hat{L}}}) $ is the inclination angle between the line of sight unit vector $\mathbf{\hat{n}} $ and the orbital angular momentum unit vector $\mathbf{\hat{L}}$. 
E.g., for an orbit seen edge on, $\iota = \pi/2$ which is actually the case of a purely plus-polarized wave, $h_\times=\cos(\pi/2)=0$.
When $\iota=0$ the orbit is seen face-on and we have a circularly-polarized wave: $\langle h_+ \rangle = \langle h_\times\rangle$ where $\langle \ldots \rangle$
denote time-averages over one orbit. Note that as $f$ increases in time, so does $h(t)$ hence the characteristic ``chirping'' of GW signals.

@@

where $M \equiv m_1+m_2 = 2.8 M_\odot$. The true ISCO is located at $r< r_\text{ISCO}$, however, there is no simple expression for it.

@@

Introducing the subscript ${}_s$ to denote quantities evaluated at the source frame, we have the following cosmological corrections

@@

@@

Given that $\dot{f}_s\sim f^{11/3}_s, h(t)\sim D^{-1} f_s^{2/3}$,
the observed frequency evolution and the GW strain change according to % in the detector frame

@@

h(t) & \sim (1+z)^{5/3} h(t).

@@

Looking at Eqs.~(\ref{eq:fdot}, \ref{eq:hplus_TD}, \ref{eq:hcross_TD}) we spot a common factor of $M_c^{5/3}$ in all of them.

@@

which changes Eqs.~(\ref{eq:fdot}, \ref{eq:hplus_TD}, \ref{eq:hcross_TD}) to

@@

h_c(t) & = \f{4}{D}\left(\f{G \mathcal{M}_c}{c^2}\right)^{5/3}\left(\frac{\pi f}{c}\right)^{2/3} \label{eq:strain_TD_redshifted}\, ,

@@

For our work here, the only important consequence is the redshifting of the frequency at which we truncate the inspiral, 
i.e., $f_\text{ISCO}\simeq 1571\,$Hz in the source frame.
For our work, we consider a maximum BNS distance of 1\,Gpc which is reasonable given the sensitivity
of A-LIGO and the ET. $D = 1\,$Gpc translates to $z \approx 0.25$.
Therefore, in the detector frame $f_\text{ISCO}$ is redshifted to $f \approx 1257\,$Hz. 
From our computations above, we know that this shift adds a negligible correction to the timescales that interest us here.
Thus we may safely disregard the effects cosmology.

@@

%Furthermore, the physical quantities that we actually observe are blind to this redshift correction. Only after high-precision measurements of
%the chirp mass and the GW polarizations $h_+, h_\times$ can we extract $z$. 
% can be obtained to a very
%high precision in the case of BNSs from the data for $\dot{f}$; e.g., $\mathcal{M}_c = 1.188^{+0.004}_{-0.002}$ for GW170817.
%The ratio $h_+/ h_\times$ gives us the orbital inclination $\iota$.
\end{enumerate}

@@

@@

\includegraphics[height=9cm]{Figures/GW170817_strains.pdf}

@@

\includegraphics[width=\linewidth]{Figures/ALigo_strains.pdf}

@@

\subsubsection{Forecasting GRBs in the 2030s using the Einstein Telescope}\label{Sec:ETB}
We now wish to quantify the forecasting capabilities of Einstein Telescope. To this end, we consider the inspiral of BNS systems
at luminosity distances of $100,200,400,1000\,$Mpc entering the ET band. The SNR accumulated at ET by each inspiral is given
by a slight modification to Eq.~(\ref{eq:ET_SNR})

%
\be
\bar\rho_{\text{tot,ET}} = \f{36}{25}A^2 h_0^2 \int_{f_0}^{f_\text{ISCO}} d f\, \f{f^{-7/3}}{S_n(f)} \label{eq:ET_SNRv2},
\ee
where $f_0$ is the GW frequency at which the emitted GWs enter ET's detection band, i.e, $f_0 < f_\text{ISCO}$ such that
%
 \be
 \sqrt{S_\text{n}(f_0)} = 2\sqrt{f_0}\, \tilde{H}_\bigtriangleup(f_0, D) \label{eq:f0_ET}\,
 \ee
 %
with $\tilde{H}_\bigtriangleup(f, D)$ given by Eq.~(\ref{eq:H_delta}).
For the detector noise $\sqrt{S_n(f)}$, we adopt the ASD for the B and C configurations known as ET-B and ET-C, respectively \cite{Hild:2010id} which we plot as the solid dark (red) and light (brown) curves in Fig.~\ref{fig:ETB2030}.
Sweeping across these are four BNS inspirals at $D=100,200,400,1000\,$Mpc represented by solid (black), dotted (blue), dashed (green),
and dot-dashed lines (gray), respectively. These strains sit much higher than the detector noise
comparred with those in Fig.~\ref{fig:LIGO2020} which foretells us that the SNRs in the ET band will be much higher, 
thus the resulting advance warning times will be much longer.

\begin{figure}[h!]
\includegraphics[width=\linewidth]{Figures/ET_strains.pdf}
\caption{Typical GW sources that may be harbingers of GRBs in the 2030s: $M=2.8 M_\odot$ inspiralling BNS systems sweeping across 
the Einstein Telescope's sensitivity band for both B and C configurations.
The solid (black), dotted (blue), dashed (green), and dot-dashed lines (gray) lines are the RMS-averaged strains, $2\sqrt{f}\tilde{H}_\bigtriangleup(f,D)$, at luminosity distances of $D=100, 200, 400, 1000\,$Mpc, respectively. 
The vertical dashed (black) line marks $f_\text{ISCO} \simeq 1571\,$Hz at which point we terminate the inspirals,
but numerical simulations indicate that the actual ISCO is smaller than the Schwarzschild value $6GM/c^2$ hence the true ISCO
frequency is greater than $f_\text{ISCO}$ used here.
As the plot shows, Einstein Telescope is sensitive enough in this regime to accumulate more SNR.
%Once again, keep in mind that we artificially extended the leading-order $f^{-2/3}$ power law past $f\sim 100\,$Hz at which point tidal effects actually visibly modify this behavior.
}
\label{fig:ETB2030}
\end{figure}
%
%
To determine $T_\text{AW}$ we once again impose the threshold SNR of 15 which is our criterion for sending out warnings to electromagnetic
observatories. Thus, rewriting Eq.~(\ref{eq:SNR_bar}) in the case of ET, we define $\bar{f}_\text{ET}$ as the solution to
%
\be
15=\f{36}{25}A^2 h_0^2 \int_{\bar{f}_\text{ET}}^{f_\text{ISCO}} d f\, \f{f^{-7/3}}{S_n(f)} .\label{eq:ET_fbar}
\ee
%
This gives us our advance warning time $T_\text{AW}=\tau_\text{insp}(\bar{f}_\text{ET})$ using Eq.~(\ref{eq:tau_insp}) which we
can slightly improve by using the 3.5PN expression. Therefore we will proceed in a manner similar to the procedure summarized in 
Fig.~\ref{fig:flow}) and compute the 0PN and 3.5PN advance warning times for all four inspirals in both the ET-B and ET-C bands.
We define the final accumulated SNR as
%
\be
\bar\rho_{F,\,\text{ET}}\equiv \f{36}{25}A^2 h_0^2 \int_{f_0}^{f_\text{ISCO}} d f\, \f{f^{-7/3}}{S_n(f)}  .\label{eq:rhoF_ET}
\ee
%

\begin{table}[h]
\centering
\begin{tabular}{lcccc}
\hline
$D\,$(Mpc) & $\bar{f}\,$(Hz) & $T^{0PN}_\text{AW}$(sec) & $T^{3.5PN}_\text{AW}$(sec)& $\bar{\rho}_{F,\,\text{ET}}$\T\B\\
\hline
100 & 1.1435 & 2909 & 2942 & 307.1 \T\\
200  \\
\hline
\end{tabular}
\caption{ET-B}\label{table:ET}
\end{table}

\section{Binary black hole neutron star systems}
Tidal disruption, minimum BH mass $5 M_\odot$. Tidal force $\sim M/r^3 \sim M^{-2}$ hence the smaller the BH the more TD there will be.
Also need highly spinning BH to decreases the ISCO radius.

>>>>

\documentclass[amsmath,amssymb,aps,floats,amsfonts,notitlepage,superscriptaddress,eqsecnum,nofootinbib,10pt]{revtex4-1}
\linespread{1.2}
%\usepackage{booktabs} % for tabular improvments
\usepackage{graphicx}       % Include figure files
\usepackage{dcolumn}        % Align table columns on decimal point
\usepackage{bm}             % bold math
\usepackage{tensor}
\usepackage{url}
\usepackage[colorlinks]{hyperref}
\usepackage{color}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\usepackage{float}
\usepackage[usestackEOL]{stackengine} % package used to write several lines in one box
\allowdisplaybreaks[1]

     % Sarp's extra packages
\usepackage{dcolumn}
	\newcolumntype{.}{D{.}{.}{13}}
	\newcolumntype{d}[1]{D{.}{.}{#1}}
\usepackage{longtable}
\usepackage{subfigure}
\usepackage{rotating}
\usepackage[small,compact]{titlesec}
%%%%%%%%%% New few lines of code are added to fix the TexLive titlesec bug in Ubuntu >= 14.04
\usepackage{etoolbox}
\makeatletter
\patchcmd{\ttlh@hang}{\parindent\z@}{\parindent\z@\leavevmode}{}{}
\patchcmd{\ttlh@hang}{\noindent}{}{}{}
\makeatother
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{fancyhdr,amsfonts}
\usepackage{times}
\usepackage{multirow}
\usepackage{sidecap}
\usepackage{eurosym}
\usepackage{setspace}
%\bibliographystyle{apsrev4-1}

\definecolor{CiteColor}{rgb}{0,0.5,0}
\hypersetup{citecolor=CiteColor}
\definecolor{RefColor}{rgb}{0.55,0,0}
\hypersetup{linkcolor=RefColor}
\definecolor{darkgreen}{rgb}{0.2,0.7,0.2}

\newcommand{\beq}{\begin{equation}}
\newcommand{\eeq}{\end{equation}}

\newcommand{\bs}{\mathbf{s}}
\newcommand{\bS}{\mathbf{S}}
\newcommand{\bp}{\mathbf{p}}
\newcommand{\bn}{\mathbf{n}}
\newcommand{\bom}{\boldsymbol{\omega}}
\newcommand{\bOm}{\boldsymbol{\Omega}}
\newcommand{\tilE}{\tilde{E}}
\newcommand{\tilL}{\tilde{L}}
\newcommand{\nn}{\nonumber}
\newcommand{\EE}{E}
\newcommand{\LL}{L}
\newcommand{\KK}{\mathcal{K}}

% Sarp's new commands
\newcommand{\ord}{\mathcal{O}}
\newcommand{\la}{\langle}
\newcommand{\ra}{\rangle}
\renewcommand{\c}{\cos}
\renewcommand{\t}{\theta}
\renewcommand{\b}{\bar}
\newcommand{\f}{\frac}
\newcommand{\pd}{\partial}
\newcommand{\bt}{\beta}
%\newcommand{\s}{\sin}
\newcommand{\ph}{\phi}
\newcommand{\g}{\gamma}
\newcommand{\al}{\alpha}
\newcommand{\La}{\Lambda}
\newcommand{\el}{\ell}
\newcommand{\h}{\hat}
\newcommand{\mrm}{\mathrm}
\newcommand{\Om}{\Omega}
\newcommand{\veps}{\varepsilon}
\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\ba}{\begin{eqnarray}}
\newcommand{\ea}{\end{eqnarray}}
\newcommand{\bi}{\begin{itemize}}
\newcommand{\ei}{\end{itemize}}
\newcommand{\bef}{\begin{frame}}
\newcommand{\ef}{\end{frame}}
\newcommand{\es}{\end{small}}
\newcommand{\Lim}[1]{\raisebox{0.5ex}{\scalebox{0.8}{$\displaystyle \lim_{#1}\;$}}}
\newcommand\T{\rule{0pt}{2.6ex}}       % Top strut for vertical spacing in tables
\newcommand\B{\rule[-1.2ex]{0pt}{0pt}} % Bottom strut

\newcommand{\Delhat}{\hat{\delta}}
\newcommand{\vsp}{\vspace{0.2cm}}

\def\D{\displaystyle}
\def\etal{\textit{et al.}}

\providecommand{\SA}[1]{{\textcolor{DarkBlue}{\texttt{SA: #1}}}}
\providecommand{\SD}[1]{{\textcolor{DarkRed}{\texttt{SD: #1}}}}

\begin{document}

\title{Forecasting Gamma-ray Bursts using Gravitational Waves}

\author{Sarp Akcay}
\affiliation{School of Mathematics \& Statistics, University College Dublin, Belfield, Dublin 4, Ireland.}

\date{\today}

\begin{abstract}
We explore the intriguing possibility of employing future gravitational-wave interferometers to detect the inspiral of compact binaries early enough to alert the electromagnetic observatories so that a gamma-ray burst (GRB) can be observed in its entirety from its very start.
We quantify the ability to predict a GRB by computing the time a canonical binary neutron star (BNS) system takes to inspiral from a certain threshold frequency to its final merger. We define this threshold frequency to be the gravitational-wave frequency at which the RMS-averaged strain due to the BNS system accumulates a signal-to-noise (SNR) ratio of 15. For our computation of advance warning times, we consider BNS systems at luminosity distances of (i) $D\lesssim200\,$Mpc in a three-interferometer network with Advance LIGO's design sensitivity and (ii) $D \lesssim 1000\,$Mpc in a similar network, but with the Einstein Telescope's B configuration
sensitivity. We find in the case of that Advanced LIGO we may get a few minutes of warning time if we get lucky whereas the Einstein Telescope will provide us
with warning times as long as a few hours.
Our main conclusions are: (i) we expect ALV network to forecast 1 to 3 GRBs in the 2020s, (ii)

\end{abstract}
\maketitle

@@

Cite the BNS Living review many times \cite{Faber:2012rw}.

@@

Let us now turn our attention to the GWs generated by the inspiral.
The GWs themselves actually are tensor distortions $h_{\mu\nu}$ of the flat Minkowski
spacetime propagating at the speed of light.
In more technical terms, they satisfy $\Box h_{\mu\nu} = 0$, i.e., the flat spacetime sourceless wave equation
meaning that we are considering the solutions far from the GW source, in the so-called wave zone defined by the condition $ c/f \ll D$.
In a suitable gauge, such as the commonly used TT (transverse-traceless) gauge, it can be shown that
there exist only two physical degrees of freedom which manifest themselves
as two independent polarization amplitudes $h_+$ and $h_\times$. %when interacting with masses away from the source.
For a circular binary at a distance $D$ these states read
%
\begin{align}
 h_+(t) &= h_c(t)\, \left(\f{1+\cos^2\iota}{2}\right)\, \cos[\Phi_\text{N}(t)],\label{eq:hplus_TD}\\
 h_\times(t) & =h_c(t)\,\cos\iota \sin[\Phi_\text{N}(t)]\label{eq:hcross_TD},
\end{align}
%
%
where
%
\be
h_c(t) =\f{4}{D}\left(\f{G M_c}{c^2}\right)^{5/3}\, \left(\f{\pi f(t)}{c}\right)^{2/3} \label{eq:h_c}
\ee
is the characteristic strain.
$\iota=\cos^{-1}(\mathbf{\hat{n}}\cdot{\mathbf{\hat{L}}}) $ is the inclination angle between the line of sight unit vector $\mathbf{\hat{n}} $ and the orbital angular momentum unit vector $\mathbf{\hat{L}}$. 
E.g., for an orbit seen edge on, $\iota = \pi/2$ which is actually the case of a purely plus-polarized wave, $h_\times=\cos(\pi/2)=0$.
When $\iota=0$ the orbit is seen face-on and we have a circularly-polarized wave: $\langle h_+ \rangle = \langle h_\times\rangle$ where $\langle \ldots \rangle$
denote time-averages over one orbit. Note that as $f$ increases in time, so does $h(t)$ hence the characteristic ``chirping'' of GW signals.

@@

where $M \equiv m_1+m_2 = 2.8 M_\odot$. The true ISCO is located at $r< r_\text{ISCO}$ with frequency possibly up to 1.7\,kHz \cite{Marronetti:2003hx}, 
however, there is no simple expression for it.

@@

Introducing the subscript $s$ to denote quantities evaluated at the source frame, we have the following cosmological corrections

@@

dt& \rightarrow (1+z) dt_s, \\

@@

Given that $\dot{f}_s=df_s/dt_s\sim f^{11/3}_s, h_s(t)\sim D^{-1} f_s^{2/3}$, and $\tau_{\text{insp},s}\sim f_s^{-8/3}$,
the observed frequency evolution, the GW strain, and the inspiral time change according to % in the detector frame

@@

h(t) & \sim (1+z)^{5/3} h(t),\\
\tau_\text{insp}& \sim (1+z)^{-5/3} \tau_{\text{insp}}

@@

Looking at Eqs.~(\ref{eq:fdot}, \ref{eq:tau_insp}, \ref{eq:h_c}) we spot a common factor of $M_c^{5/3}$ in all of them.

@@

which changes Eqs.~(\ref{eq:fdot}, \ref{eq:tau_insp}, \ref{eq:h_c}) to

@@

h_c(t) & = \f{4}{D}\left(\f{G \mathcal{M}_c}{c^2}\right)^{5/3}\left(\frac{\pi f}{c}\right)^{2/3} \label{eq:strain_TD_redshifted}\\
\tau_\text{insp}(f) &= \f{5}{256\pi}\f{c^5}{(\pi G \mathcal{M}_c)^{5/3}} \,f^{-8/3}\label{eq:tau_insp_redshifted}\, ,

@@

Therefore, the inspiral evolves faster in the lab frame %because $\dot{f}$ has increased and the cut-off frequency has decreased.
and we must now terminate it at a redshifted GW ISCO frequency which is less than $f_\text{ISCO}$ of Eq.~(\ref{eq:f_isco}).
On the other hand, the Fourier transform of $h_c(t)$ changes to 

@@

\tilde{H}(f) = A\, \f{c}{D} \f{1}{(1+z)} \left(\f{G \mathcal{M}_c}{c^3}\right)^{5/6} f^{-7/6}
indicating that the frequency-domain strain in the detector frame decreases by a factor of $(1+z)^{-1/6}$. %\approx 1 -z/6$ for small $z$.

@@

For the Advanced-LIGO-Virgo network,
we consider a maximum BNS luminosity distance of 200\,Mpc, 
i.e., $z \approx 0.022$ which redshifts $f_\text{ISCO}$ by $\approx 2\%$ and reduces $\tau_\text{insp}$ by $\approx 7\%$.
Therefore, we will neglect the effects of cosmological redshifting on the less sensitive ALV network.
On the other hand, for Einstein Telescope, we will go out to $D=1\,\text{Gpc}\implies z\approx 0.2$
which reduces $\tau_\text{insp}$ by $1/4$, thus we shall include the redshift effects in our computations of advance warning times and SNRs
for which we provide the details in Sec.~\ref{Sec:ETB} (additional details can be found in, e.g., Ref.~\cite{Phinney:2001di}).

%
%Furthermore, the physical quantities that we actually observe are blind to this redshift correction. Only after high-precision measurements of
%the chirp mass and the GW polarizations $h_+, h_\times$ can we extract $z$. 
% can be obtained to a very
%high precision in the case of BNSs from the data for $\dot{f}$; e.g., $\mathcal{M}_c = 1.188^{+0.004}_{-0.002}$ for GW170817.
%The ratio $h_+/ h_\times$ gives us the orbital inclination $\iota$.
\end{enumerate}
%
%

@@

\includegraphics[height=9cm]{../Figures/GW170817_strains.pdf}

@@

\includegraphics[width=\linewidth]{../Figures/ALigo_strains.pdf}

@@

\subsubsection{Forecasting GRBs in the 2030s using the Einstein Telescope}\label{Sec:ETB}
We now wish to quantify the forecasting capabilities of Einstein Telescope. To this end, we consider the inspiral of BNS systems
at luminosity distances of $100,200,400,1000\,$Mpc entering the ET band. 
These correspond to cosmological redshifts of $z\approx 0.0222,0.0437, 0.085, 0.198$, respectively.
To compute these, we used a flat $\Lambda$CDM model ($\Omega_k=0$) with the latest Planck satellite parameters: 
$\Omega_\Lambda = 0.6911, \Omega_m = 0.3089, H_0 = 67.74\,$km\,s$^{-1}\,$Mpc$^{-1}$ \cite{Planck2015}. % with negligible ra yielding and $\Omega_r \lesssim 10^{-4}$.
It is then straightforward to translate $D$ to $z$ (cf. Ref.~\cite{Hogg:1999ad}).

The SNR accumulated in ET from each inspiral is obtained by a modification to Eq.~(\ref{eq:ET_SNR})

%
\be
\bar\rho_{\text{ET}}(f_0,f)\equiv \f{36}{25}A^2 h_0^2\, (1+z)^{-1/3}\int_{f_0}^{f} d f'\, \f{f'^{-7/3}}{S^\text{ET}_n(f')} \label{eq:ET_SNRv2},
\ee
where $f_0$ is the GW frequency at which the emitted GWs enter ET's detection band, i.e, $f_0 < f_\text{ISCO}$ such that
%
 \be
 \sqrt{S^\text{ET}_\text{n}(f_0)} = 2\sqrt{f_0}\, \tilde{H}_\text{ET}(f_0, D) \label{eq:f0_ET}\,
 \ee
 %
with $\tilde{H}_\text{ET}(f, D)$ given by the redshifted version of Eq.~(\ref{eq:H_delta})
%
\be
\tilde{H}_\text{ET}(f, D) = \f{1}{2}\sqrt{\f{3}{10}}\, \f{c}{D}\f{1}{(1+z)}\left(\f{G \mathcal{M}_c}{c^3}\right)^{5/6} f^{-7/6}\, , \label{eq:H_delta_redshifted}
\ee
%
where recall $\mathcal{M}_c = (1+z)M_c$.
For the detector noise $\sqrt{S^\text{ET}_n(f)}$, we adopt the ASD for the B and C configurations known as ET-B and ET-C, respectively \cite{Hild:2010id} which we plot as the solid dark (red) and light (brown) curves in Fig.~\ref{fig:ETB2030}.
Sweeping across these are four BNS inspirals at $D=100,200,400,1000\,$Mpc represented by solid (black), dotted (blue), dashed (green),
and dot-dashed lines (gray), respectively. These strains sit much higher than the detector noise
compared with those in Fig.~\ref{fig:LIGO2020} which foretells us that the SNRs accumulated in the ET band will be much higher, 
thus yielding much longer advance warning times as we show in Table~\ref{table:ET}.
%As shown, Einstein Telescope is sensitive enough to accumulate more SNR in this $f>f_\text{ISCO}$ regime.

\begin{figure}[h!]
\includegraphics[width=\linewidth]{../Figures/ET_strains_redshifted.pdf}
\caption{Typical GW sources that may be harbingers of GRBs in the 2030s: $1.4 M_\odot-1.4 M_\odot$ inspiralling BNS systems sweeping across 
the Einstein Telescope's sensitivity band for both B and C configurations.
The solid (black), dotted (blue), dashed (green), and dot-dashed lines (gray) lines are the redshift-corrected
RMS-averaged strains, $2\sqrt{f}\tilde{H}_\text{ET}(f,D)$, at luminosity distances of $D=100, 200, 400, 1000\,$Mpc, respectively. 
The vertical lines with correspondingly identical patterns (colors) mark the redshifted ISCO frequencies ($(1+z)^{-1} f_\text{ISCO}$) at which point we terminate each inspiral.
As the true ISCO frequency is likely larger than $f_\text{ISCO}$ \cite{Marronetti:2003hx}, the inspirals would continue to nearly 2\,kHz indicated by the 
faded lines in the plot (drawn to 5\,kHz for aesthetic reasons).
%However, numerical simulations indicate that the actual ISCO is smaller than the Schwarzschild value $6GM/c^2$ hence the true ISCO
%frequency is greater than $f_\text{ISCO}$ used here.
%As shown, Einstein Telescope is sensitive enough to accumulate more SNR in this $f>f_\text{ISCO}$ regime.
%Once again, keep in mind that we artificially extended the leading-order $f^{-2/3}$ power law past $f\sim 100\,$Hz at which point tidal effects actually visibly modify this behavior.
}
\label{fig:ETB2030}
\end{figure}
%
%
To determine $T_\text{AW}$ we once again impose the threshold SNR of 15 which, the reader may recall, 
is our detection criterion for sending out warnings to electromagnetic observatories. Thus, rewriting Eq.~(\ref{eq:SNR_bar}) in the case of ET, we define $\bar{f}_\text{ET}$ as the solution to
%
\be
\bar\rho_\text{ET}(f_0,\bar{f}_\text{ET}) = 15 \label{eq:ET_fbar},
\ee
%
where $f_\text{ISCO}$ is now the redshifted version of Eq.~(\ref{eq:f_isco}), equalling $1537, 1505, 1448, 1311\,$Hz, respectively.
$\bar{f}_\text{ET}$ then gives us our redshift-accounted advance warning time via $T_\text{AW}=\tau_\text{insp}(\bar{f}_\text{ET})$ using Eq.~(\ref{eq:tau_insp_redshifted}) which we can slightly improve by using the 3.5PN expression. 
Therefore, we proceed as summarized in Fig.~\ref{fig:flow} and compute the 0PN and 3.5PN advance warning times 
for all four inspirals in both the ET-B and ET-C bands.
The latter computation involves rescaling the PN parameters $x$ and $\Theta$ to $(1+z)^{2/3} x$ and $(1+z)^{-1}\Theta $ (see Sec.~9.3 of 
Blanchet's review \cite{Blanchet_LRR} for details).
The final accumulated SNR is given by
%
\be
\bar\rho_{F}\equiv \bar\rho_\text{ET}(f_0,f_\text{ISCO})  .\label{eq:rhoF_ET}
\ee
%
Table~\ref{table:ET} summarizes our results for both ET-B and ET-C where we chose to only display the 3.5PN-accurate inspiral time as $T_\text{AW}$,
but we computed this quantity via both 0PN and 2PN expressions as a check.

\begin{table}[h]
\centering
\begin{tabular}{l|ccccccc}
%toprule
\hline
$D\,$(Mpc) & \multicolumn{3}{c}{ET-B} &  & \multicolumn{3}{c}{ET-C}\T\B\\
\hline
{}& $\bar{f}\,$(Hz) & \ $T_\text{AW}$(sec) \ & $\bar{\rho}_{F}$ &{} & $\bar{f}\,$(Hz) & \ $T_\text{AW}$(sec)\ & $\bar{\rho}_{F}$\T\B\\
%midrule

100 & 6.7237 &  2818 & 306.0 &{\qquad} & 3.2692 & 19212 & 365.4\T\\
200 & 11.2104 & 698.6 & 152.4 &{\qquad} & 4.1028 & 10138 & 182.1 \\
400 & 18.2234 & 179.8 & 75.74 &{\qquad} & 5.0607 & 5437 & 90.48\\
1000 & 41.3537 &17.2 & 29.79& \qquad & 6.7553 & 2138 & 35.60  \\
\hline
%bottomrule
\end{tabular}
\caption{Forecasting capabilities of Einstein Telescope summarized. We can expect over \emph{five} hours of early warning for nearby sources. 
ET-B and ET-C refer the different configurations shown in Fig.~\ref{fig:ETB2030}. For the advance warning times, we only present the result of the more accurate 3.5PN computation. $\bar{f}$ is the threshold frequency at which
ET-B/C accumulate SNR of 15 which we take to be our detection criterion. Note that both $T_\text{AW}$ and $\bar\rho_F$ are larger for ET-C
due to its improved sensitivity in the $f\lesssim 30\,$Hz regime compared to ET-B as it is clear in Fig.~\ref{fig:ETB2030}.
These results and those of Table.~\ref{table:LIGO2020} are summarized in Fig.~\ref{fig:summary}.}\label{table:ET}
\end{table}
%
%
As the results of Table~\ref{table:ET} indicate, ET-B will provide $\sim 10 - 50\,$ minutes of warning time for BNS inspirals at $D\lesssim 200\,$Mpc.
Instead of speculating whether or not this is enough to forecast GRBs, we turn our attention to ET-C's forecasting capabilities.
As can be seen from Fig.~\ref{fig:ETB2030}, ET-C's increased sensitivity in the $f\lesssim 30\,$Hz regime will significantly lengthen the advance warning times.
This is best highlighted by the sixth colum of Table~\ref{table:ET}, 
where we show that ET-C will provide up to five hours of early warning time for nearby ($D\lesssim 100\,$Mpc) neutron star mergers.
Moreover, BNS inspirals as far as 200\,Mpc (the ALV network's horizon distance) will persist in the ET-C sensitivity band for a few hours after their initial detection.
Current BNS merger rate, inferred from A-LIGO's O1-O2 periods, implies that $\sim 40$ to $\sim600$ sources within a volume of $(200\,\text{Mpc})^3$ will be detected yearly by ET.
Speculating that the advances in the detection, localization and early-warning algorithms in the next decade will reduce the required early
warning time to less than an hour, we set $T_\text{AW} = 1\,\text{hour}$ and obtain corresponding BNS horizon distances of $D_H=87,613\,$Mpc ($z\approx 0.02,0.127$) for ET-B and ET-C, respectively.
In other words, all inspiralling BNSs closer than $D_H$ will give us more than an hour's leeway before the GRB (assuming detection SNR of 15 as always).
The volume set by $D=D_H$ translates to a BNS merger rate of $\mathcal{O}(1)$ for ET-B and
$\approx 355^{+730}_{-280}\,\text{year}^{-1}$ for ET-C. Additionally, each source within $D_H$ will accumulate $\bar\rho_F \gtrsim 420, 58$ over its inspiral
in the ET-B, ET-C bands, respectively. These are summarized in Table~\ref{table:horizon}.
